---
title: "datawrangling0930"
output: Github_document
editor_options: 
  chunk_output_type: inline
---
# Rules for tidy data

- observations(row) x variables (col)
- every *value* has a cell

- relational data: spread across tables w/ defined relations
  - keys: variables used to define these relations
  - tables are combined by joins: left, right, full
    - L and R are the same, stick to 1 when using
    
- key functions:in `tidyr` `dplyr`
  - `pivot_longer` like `gather`, but `gather` is confusing
  - `separate`
  - `pivot_wider` less used, like `spread`, but `spread` is confusing
  - `bind_rows`
  - `*_join` left/right/full
  
# practice
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## `pivot_longer`

- wide format to long format
- value in variables --> value in cells 
- e.g., public_pulse_data

```{r}
library(tidyverse)

pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

view(pulse_df)

# need to put f/u time into cells

pulse_tidy =
  pivot_longer(
    pulse_df,
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", # variable name to the f/u time
    names_prefix = "bdi_score_", # delete the prefixes
    values_to = "bdi" # values of the original variable
  ) %>%

relocate(visit) %>%
mutate(
    visit = replace(visit, visit == "bl", "00m"),#change baseline into 0 months
    visit = factor(visit) # let the visit time transformed into factor
    ) %>%

arrange(id, visit)
view(pulse_tidy)
```

## `pivot_wider`

- long format into wide format

```{r}
# data to be used

analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

# not reading-friendly in reading, transpose time into columns

analysis_result = 
  pivot_wider(
    analysis_result,
    names_from = "time",
    values_from = "mean"
  )
view(analysis_result)

```

- ambiguity for tidiness
  - according to the usage
  
## `bind_row`

```{r}
# put 3 tables of different movies into 1 table

fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring") # add a variable "movie"

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")

view(return_king)

## merge the data

lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>% # put the rows together (more rows)
  janitor::clean_names() %>%
  pivot_longer(
    female:male, # the cols to be changed
    names_to = "sex", # the new variables for cols
    values_to = "words" # the new variables for the previous values in "male" and "female"
  ) %>%
  mutate(race = str_to_lower(race)) %>%
  select(movie, everything())

view(lotr_tidy)
```

- rbind vs. bind_rows
  - `rbind` is problematic
  - never use `rbind`


## `*_join`

- left? right? full? inner?
- default for `*_join`: R will try to detemine the keys based on he variable names
- `*_join` cannot be followed by `%>%`: join is not a linear step

```{r}
# join pups and litters

pup_data = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(
    sex = recode(sex, `1` = "male", `2` = "female"),
    sex = factor(sex)) 

litter_data = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>% # the groups are not looking good, thus separate the group into 2 variables, separate them after the 3rd character
  relocate(litter_number) %>% # litter_number is the key variable, put it into the leading column
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    dose = str_to_lower(dose))

view(litter_data)

# merge data by litter_number (key variable)
fas_data = 
  left_join(pup_data,litter_data, by = "litter_number")

view(fas_data)

```

# questions

- knit output table for regression?
  - latex?
  - `roomtidy`?

